---
title: "Steps Analysis"
author: "P. Mae Cooper"
date: "November 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r results='hide'}
library(lubridate, quietly = TRUE)
library(stringr, quietly = TRUE)

#get data
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
activity<- read.csv(unz(temp, "activity.csv"))
unlink(temp)

#format dates and times
activity$date<- ymd(activity$date)

activity$interval<- str_pad(activity$interval, 4, pad = "0")
activity$interval<- hm(paste(substr(activity$interval,1,2),":",substr(activity$interval,3,4)))

#create version without missing days
activity2<- subset(activity,is.na(activity$steps)==FALSE)
```
  
This is an analysis of the number of steps taken throughout the day, based single person's activity. The data cover 53 days between October 2nd and November 29th, 2012.There was no data recorded on October 8th, or November 1st, 4th, 9th, 10th, or 14th.  
  
First, we will examine total steps in a day. The distribution of steps per day looks like this:

```{r}
library(ggplot2, quietly = TRUE)

#summarize steps per day
daysteps<- with(activity2,tapply(steps,date,sum))

#plot
qplot(daysteps, main="Total Daily Steps", ylab = "Days", xlab = "Steps",
      bins=25, fill=I("dark green"), col=I("gray"))

#get average statistics
mn<- round(mean(daysteps),0)
md<- median(daysteps)
```
  
The number of daily steps has a mean of `r format(mn, scientific=FALSE, big.mark = ",")` and a median of `r format(md,big.mark = ",")`.  
  
Next, the average number of steps over the course of a day:  
  
```{r}
library(plyr, quietly = TRUE)
library(dplyr, quietly = TRUE)


#summarize steps per interval - mean and standard deviation
timesteps<- activity2 %>% 
            group_by(as.character(interval)) %>% 
            summarize(mn_steps=mean(steps),std_step=sd(steps)) %>% 
            rename(interval_chr="as.character(interval)") %>%
            mutate(interval=case_when(
                  is.na(hms(interval_chr)) & is.na(ms(interval_chr)) ~ hms("00:00:00"),
                  is.na(hms(interval_chr)) & !is.na(ms(interval_chr)) ~ ms(interval_chr),
                  TRUE ~ hms(interval_chr)))%>%
            arrange(as_datetime(interval))

#plot
plot(as_datetime(timesteps$interval),timesteps$mn_steps, type='l',
     main="Average Steps Across the Day", ylab = "Steps", xlab = "Time",
     col="red", lwd=2.5)

#get max stats
max<- max(timesteps$mn_steps)
timeMax<- as_datetime(timesteps$interval)[timesteps$mn_steps==max]
```
  
On average, steps peak at `r format(timeMax,"%H:%M")` with `r format(max,digits=0)` steps.

```{r results='hide'}
#number of rows with missing values
NA_rows<- length(activity$steps[is.na(activity$steps)])

#create variables for merging
timesteps$mg<- as_datetime(timesteps$interval)
activity$mg<-  as_datetime(activity$interval)

#merge activity data with summary data
activity3<-merge(activity, timesteps, by="mg",all=TRUE)
activity3<- activity3%>%
      select(steps,date,interval.x,mn_steps,std_step)%>%
      arrange(date,as_datetime(interval.x))%>%

#simulate missing values using mean of each time interval with poisson distribution
mutate(new_steps=ifelse(!is.na(steps),steps,rpois(10000,lambda=mn_steps)))

#get average statistics, including imputed data
daysteps2<- with(activity3,tapply(new_steps,date,sum))
mn2<- round(mean(daysteps2),0)
md2<- median(daysteps2)
```
With the missing days, there are `r NA_rows` rows with missing values.  
  
We can impute the missing data by picking from a poisson distribution based on the 
average number of steps for each interval. If we do so, the distribution of steps per day looks like this:

```{r}
qplot(daysteps2, main="Total Daily Steps (with imputed data)", ylab = "Days", xlab = "Steps",
      bins=25, fill=I("dark green"), col=I("gray"))
```

We can see that this increases the number of days that are close to the center of the distribution, since the imputed days are based on mean tendencies. The new mean, `r format(mn, scientific=FALSE, big.mark = ",")`, and median, `r format(md,big.mark = ",")`, are fairly similar to the old mean and median. It is only the variation that has decreased.

